### 2021-09-06 paper/header.tex
add initial intro draft

% **************GENERATED FILE, DO NOT EDIT**************

\title{ProbabilityBoundsAnalysis.jl:\\ an arithmetic of probabilities and sets of probabilities}

\author[1]{Ander Gray}
\author[1]{Scott Ferson}
\author[1, 2]{Edoardo Patelli}
\affil[1]{Institute for Risk and Uncertainty, University of Liverpool}
\affil[2]{Centre for Intelligent Infrastructure, University of Strathclyde}

\keywords{Julia, Probability, Uncertainty, Interval Arithmetic, Probabilistic Arithmetic}

\hypersetup{
pdftitle = {ProbabilityBoundsAnalysis.jl:\\ an arithmetic of probabilities and sets of probabilities},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Ander Gray, Scott Ferson, Edoardo Patelli},
pdfkeywords = {Julia, Probability, Uncertainty, Interval Arithmetic, Probabilistic Arithmetic},
}